---
title: "Formelsammlung"
author: "Silvan Hüsler"
header-includes:
   - \usepackage{bm}
   - \usepackage{amsmath,amssymb,mathrsfs,booktabs}
   - \newcommand{\matr}[1]{\mathbf{#1}} % undergraduate algebra version
output: 
  pdf_document:
    keep_tex: false
    highlight: pygments
    number_sections: false
nocite: | 
  @Durrant.2013, @West.2014, @James.2013, @Steele.2005
references:
- id: Durrant.2010
  type: article-journal
  author:
  - family: Durrant
    given: Gabriele B.
  - family: Groves
    given: Robert M.
  - family: Staetsky
    given: Laura
  - family: Steele
    given: Fiona
  issued:
  - year: '2010'
  title: EFFECTS of interviewer attitudes and behaviors on refusal in household surveys
  container-title: The Public Opinion Quarterly
  page: '1-36'
  volume: '74'
  issue: '1'
  abstract: Interviewers play a crucial role in gaining cooperation from a sample
    unit. This paper aims to identity the interviewer characteristics that influence
    survey cooperation. Of principal interest to survey practitioners are interviewer
    attributes associated with higher cooperation rates, particularly among sample
    members whose characteristics are traditionally associated with a lower probability
    of response. Our data source is unusually rich, in that it contains extensive
    information on interviewers, including their attitudes and behaviors, which are
    linked to detailed information on both responding and nonresponding sample units.
    An important value of the data is that they permit examining a host of as yet
    unanswered questions about whether some interviewer attributes stimulate cooperation
    among some respondents but not others. In short, we investigate whether some sample
    units react favorably to certain interviewer characteristics. A multilevel cross-classified
    logistic model with random interviewer effects is used to account for clustering
    of households within interviewers, due to unmeasured interviewer attributes, and
    for the cross-classification of interviewers within areas. The model allows for
    statistical interactions between interviewer and household characteristics.
  ISSN: 0033362X, 15375331

- id: Durrant.2013
  type: chapter
  author:
  - family: Durrant
    given: Gabriele B.
  - family: D'arrigo
    given: Julia
  - family: Müller
    given: Gerrit
  editor:
  - family: Kreuter
    given: Frauke
  issued:
  - year: '2013'
  title: 'Modeling call record data: Examples from cross-sectional and longitudinal
    surveys'
  title-short: Modeling call record data
  container-title: Improving surveys with paradata
  publisher: John Wiley & Sons, Inc
  publisher-place: Hoboken, New Jersey
  page: '281-308'
  DOI: 10.1002/9781118596869.ch12
  ISBN: '9781118596869'

- id: West.2014
  type: book
  author:
  - family: West
    given: Brady T.
  - family: Welch
    given: Kathleen B.
  - family: Galecki
    given: Andrzej T.
  issued:
  - year: '2014'
  title: 'Linear mixed models: A practical guide using statistical software, second
    edition'
  title-short: Linear mixed models
  publisher: Taylor and Francis
  publisher-place: Hoboken
  number-of-pages: '434'
  edition: 2\. ed.
  abstract: 'Highly recommended by JASA, Technometrics, and other journals, the
    first edition of this bestseller showed how to easily perform complex linear mixed
    model (LMM) analyses via a variety of software programs. Linear Mixed Models:
    A Practical Guide Using Statistical Software, Second Edition continues to lead
    readers step by step through the process of fitting LMMs. This second edition
    covers additional topics on the application of LMMs that are valuable for data
    analysts in all fields. It also updates the case studies using the latest versions
    of the software procedures and provides up-to-date in'
  ISBN: '978-1-4665-6102-1'

- id: James.2013
  type: book
  author:
  - family: James
    given: Gareth
  - family: Witten
    given: Daniela
  - family: Hastie
    given: Trevor
  - family: Tibshirani
    given: Robert
  issued:
  - year: '2013'
  title: An introduction to statistical learning
  publisher: Springer New York
  URL: http://dx.doi.org/10.1007/978-1-4614-7138-7
  DOI: 10.1007/978-1-4614-7138-7

- id: Steele.2005
  type: no-type
  author:
  - family: Steele
    given: F.
  issued:
  - year: '2005'
  title: Event history analysis
  publisher: NCRM/004
  publisher-place: NCRM Methods Review Papers
  abstract: 'Highly recommended by JASA, Technometrics, and other journals, the
    first edition of this bestseller showed how to easily perform complex linear mixed
    model (LMM) analyses via a variety of software programs. Linear Mixed Models:
    A Practical Guide Using Statistical Software, Second Edition continues to lead
    readers step by step through the process of fitting LMMs. This second edition
    covers additional topics on the application of LMMs that are valuable for data
    analysts in all fields. It also updates the case studies using the latest versions
    of the software procedures and provides up-to-date in'
---


# A James et al. 2013: Logistic Regression

\begin{align}
\text{Pr}(\text{default} &= \text{yes}|\text{balance}) \notag \\
\text{p}(X) &= \beta_0 + \beta_1 X \tag{4.1} \\
\text{p}(X) &= \frac{e^{\beta_0 + \beta_1 X}}{1-e^{\beta_0 + \beta_1 X}} \tag{4.2: logistic function} \\
\frac{\text{p}(X)}{1-\text{p}(X)}  &= e^{\beta_0 + \beta_1 X} \tag{4.3} \\
&=\frac{\text{p}(X)}{1-\text{p}(X)} \tag{odds} \\
\text{log} \left(\frac{\text{p}(X)}{1-\text{p}(X)} \right)  &= \beta_0 + \beta_1 X \tag{4.4: lhs -> log-odds oder logit}
\intertext{Wie wirkt sich eine Einheit mehr von X aus?}
&\text{-> } \frac{\text{p}(X)}{1-\text{p}(X)} \times e^{b_1} \notag
\end{align}

<!-- *+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*  -->

# B West et al. 2014: Linear Mixed Models



## 3 Two-Level Models for Clustered Data: The Rat Pup Example

### A General Model

\begin{align*}
\text{WEIGHT}_{ij} = \beta_{0} + \beta_{1} &\times \text{TREAT1}_j  + \beta_{2} \text{TREAT2}_j + \beta_{3} \times \text{SEX1}_{ij} + \\
\beta_{4} &\times \text{LITSIZE}_j + \beta_{5} \times \text{TREAT1}_j \times \text{SEX1}_{ij} + \\
\beta_{6} &\times \text{TREAT2}_j \times \text{SEX1}_{ij} + \\
\\
&\text{u}_j + \varepsilon_{ij}  \\
\end{align*}




### Level 1 Model (Rat Pup)

$$
\text{WEIGHT}_{ij} = \beta_{0j} + \beta_{1j} \times \text{SEX1}_{ij} + \varepsilon_{ij}
$$


### Level 2 Model (Litter)

\begin{align*}
\beta_{0j} &= \beta_{0} + \beta_{1}  \times \text{TREAT1}_j  + \beta_{2} \text{TREAT1}_j + \beta_{4} \text{LITSIZE}_j + \text{u}_j \\
\beta_{1j} &= \beta_{3} + \beta_{5}  \times \text{TREAT1}_j  + \beta_{6} \text{TREAT2}_j 
\end{align*}



## 6 Random Coefficient Models for Longitudinal Data: The Autism Example

### A General Model

\begin{align*}
\text{VSAE}_{ti} = \beta_0 + \beta_1 &\times \text{AGE2}_{ti} + 
\\ \beta_2 &\times \text{AGE2SQ}_{ti} + 
\\ \beta_3 &\times \text{SICDEGP1}_i + 
\\ \beta_4 &\times \text{SICDEGP2}_i + 
\\ \beta_5 &\times \text{SICDEGP1}_i \times \text{AGE2}_{ti} +
\\ \beta_6 &\times \text{AGE2}_{ti} \times \text{SICDEGP2}_i \times \beta_7 \times \text{AGE2SQ} \times \text{SICDEGP1}_i + 
\\ \beta_8 &\times \text{AGE2SQ} \times \text{SICDEGP2}_i + \\
\\
\text{u}_{0i} +
\text{u}_{1i} &\times \text{AGE2}_{ti} +
\text{u}_{2i} \times \text{AGE2SQ}_{ti} +  \varepsilon_{ti}
\end{align*}




### Level 1 Model (Time)

\[
\text{VSAE}_{ti} = \beta_{0i} + \beta_{1i} \times \text{AGE2}_{ti} + \beta_{2i} \text{AGE2SQ}_{ti} + \varepsilon_{ti}
\]


### Level 2 Model (Child)



\begin{align*}
\beta_{0i} &= \beta_0 + \beta_3 \times \text{SICDEGP1}_i + \beta_4 \times \text{SICDEGP2}_i + \text{u}_{0i}\\
\beta_{1i} &= \beta_1 + \beta_5 \times \text{SICDEGP1}_i + \beta_6 \times \text{SICDEGP2}_i + \text{u}_{1i}\\
\beta_{2i} &= \beta_2 + \beta_7 \times \text{SICDEGP1}_i + \beta_8 \times \text{SICDEGP2}_i + \text{u}_{2i}\\
\end{align*}



```{r, include=FALSE}
#setwd("C:/Users/shs/Desktop")
library("tidyverse")
library("modelr")


library(AzureML)
ws <- workspace()

autism <- download.datasets(
  dataset = ws, 
  name    = "autism.csv")
attach(autism)

sicdegp.f <- factor(sicdegp)
age.f <- factor(age)
autism.updated <- data.frame(autism, sicdegp.f, age.f)

summary(age.f)
table(sicdegp.f, age.f)
summary(vsae)
tapply(vsae, age.f, mean, na.rm=TRUE)
tapply(vsae, age.f, min, na.rm=TRUE)
tapply(vsae, age.f, max, na.rm=TRUE)

age.2 <- age - 2
age.2sq <- age.2*age.2
sicdegp2 <- sicdegp
sicdegp2[sicdegp == 3] <- 0
sicdegp2[sicdegp == 2] <- 2 
sicdegp2[sicdegp == 1] <- 1
sicdegp2.f <- factor(sicdegp2)

autism.updated <- subset(data.frame(autism, sicdegp2.f, age.2), !is.na(vsae))
```

### Merke: Wichtiger Aspekt bezüglich dem Verständnis der Modelle ist Kenntnis, wie Formeln den Datensatz transformieren!

```{r}
form <- vsae ~ 
        age.2 + 
        I(age.2^2) + 
        sicdegp2.f + 
        age.2*sicdegp2.f + 
        I(age.2^2)*sicdegp2.f + 
        (age.2 + I(age.2^2) | childid)
```


```{r}
model_matrix(autism.updated, form) %>% glimpse
```


<!-- *+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*  -->

# C Steele 2005: Event History Analysis

## 2. Descriptive Event History Analysis
###  2.1 Definitions: hazard and survivor functions
###  2.2 Estimation of the hazard and survivor functions 

<!-- Seite 7 -->

\begin{align}
\hat{h}_{t} &= \frac{d_{t}}{r_{t}-w_{t}/2} \tag{2.1} \\
\hat{S}_{1}       &= 1 \notag \\
\hat{S}_{t}       &= (1-\hat{h}_{1})(1-\hat{h}_{2})\ldots(1-\hat{h}_{t-1}) \text{ for } t>1 \tag{2.2} \\
            &= \prod_{j=1}^{t-1}(1-\hat{h}_{j}) \notag
\end{align}





## 3. Continuous-time Event History Models
###  3.1 Introduction
###  3.2 Proportional hazards and accelerated life models
####  3.2.1 Proportional hazards model

\begin{align}
h(t;\matr{x}_{i})                             &= h_{0}(t)g(\matr{x}_{i}) \tag{3.1} \\
\frac{h(t;\matr{x}_{1})}{h(t;\matr{x}_{2})} &= \frac{g(\matr{x}_{1})}{g(\matr{x}_{2})} \notag
\end{align}


###  3.3 Cox proportional hazards model

\begin{align}
h(t;\matr{x}_{i})      &= h_{0}(t)\exp\left(\bm{\beta}' \matr{x}_{i}\right) \tag{3.2} \\
\log h(t;\matr{x}_{i}) &= \log h_{0}(t)+\bm{\beta}' \matr{x}_{i} \tag{3.3}
\end{align}


## 4. Discrete-time Event History Models
###  4.1 The discrete-time approach
###  4.2 Data structure for a discrete-time analysis

$y_{ti}=
\left\{\begin{array}{ll}
0 & t<y_{i}\\
0 & t=y_{i},\ \delta_{i}=1\\
1 & t=y_{i},\ \delta_{i}=0.
\end{array}\right.$

##  4.3 Discrete-time models


\begin{align}
\text{logit}(h_{ti})= \log\left(\frac{h_{ti}}{1-h_{ti}}\right)=\alpha(t)+\bm{\beta}' \matr{x}_{ti} \tag{4.1}
\end{align}



## 5. Unobserved Heterogeneity	
### 5.1 The problem and implications of unobserved heterogeneity
###  5.2 Incorporating unobserved heterogeneity
#### 5.2.1 The Cox model with frailty


\begin{align}
h(t;\matr{x}_{i}, u_i)      &= h_{0}(t) u_i \exp\left(\bm{\beta}' \matr{x}_{i}\right) \tag{5.1}
\end{align}


####  5.2.2 The discrete-time logit model with frailty
\begin{align}
\text{logit}(h_{ti})        &= \alpha(t)+\bm{\beta}' \matr{x}_{ti} + u_i \tag{5.2} 
\end{align}

###  5.3 Example: age at first partnership
## 6. Repeated Events
###  6.1 Examples
###  6.2 Issues in the analysis of repeated events
###  6.3 A discrete-time model for repeated events

\begin{align}
\text{logit}(h_{tij})          &= \alpha(t) + \bm{\beta}' \matr{x}_{tij} + u_i \tag{6.1}
\end{align}

## 7. Competing Risks
###  7.1 Introduction
###  7.2 Definitions

$h^{(r)}(t)=\displaystyle \lim_{dt\rightarrow 0}\frac{\mathrm{P}\mathrm{r}(t\leq T<t+dt,R=r|T\geq t)}{dt}.$

###  7.3 Models for competing risks
####  7.3.1 Separate models for each type of transition

$\delta_{i}^{(r)}=\left\{\begin{array}{ll}
1 & \mathrm{i}\mathrm{f}\ r_{i}\neq r\\
0 & \mathrm{i}\mathrm{f}\ r_{i}=r.
\end{array}\right.$


\begin{align}
\log h^{(r)}(t;\matr{x}_{i}^{(r)}(t))=\log h_{0}^{(r)}(t)+\bm{\beta}^{(r)} \mathrm{'} \matr{x}_{i}^{(r)}(t),  r=1,\cdots, k \tag{7.1}
\end{align}


$y_{ti}^{(r)}=
\left\{\begin{array}{ll}
0 & t<y_{i}\\
0 & t=y_{i},\ r_{i}=r\\
1 & t=y_{i},\ r_{i}\neq r.
\end{array}\right.$


\begin{align}
h_{ti}^{(r)} &= \mathrm{P}\mathrm{r}(y_{ti}^{(r)}=1|\mathrm{y}_{si}=0,s<t) \notag \\
logit (h_{ti}^{(r)}) &= \alpha^{(r)}(t)+\bm{\beta}^{(r)} \mathrm{'} \matr{x}_{ti}^{(r)}, r=1,\cdots,k \tag{7.2}
\end{align}


####  7.3.2 Modelling event types simultaneously: the multinomial logit model 

\begin{align}
h_{ti}^{(0)}                            &= \mathrm{P}\mathrm{r}(y_{ti}=0|y_{si}=0,\ s<t)=1-\sum_{r=1}^{k}h_{ti}^{(r)} \tag{7.3} \\
\log\left(\frac{h_{ti}^{(r)}}{h_{ti}^{(0)}}\right) &= \alpha^{(r)}(t)+\bm{\beta}^{(r)} \mathrm{'} \matr{x}_{ti}^{(r)} \quad r=1,\cdots,k. \tag{7.4} \\
\notag \\
h_{ti}^{(r)}                            &= \frac{\exp[\alpha^{(r)}(t)+\beta^{(r)_{1}}\mathrm{x}_{ti}^{(r)}]}
{1+\sum_{lr=1}^{k}\exp[\alpha^{(l)}(t)+\bm{\beta}^{(l)_{1}}\matr{x}_{ti}^{(l)}]}, \quad r=1, \cdots k, \tag{7.5} \\
\notag \\
\hat{S}_{1i}=1; \hat{S}_{ti} &= \prod_{j=1}^{t-1}(1-\hat{h}_{ji}) , t>1, \tag{7.6} \\
\text{where } \hat{h}_{ji} &= \sum_{r=1}^{k}\hat{h}_{ji}^{(r)}. \notag \\
\notag \\
\hat{f}_{ti}^{(r)} &=\hat{h}_{ti}^{(r)}\hat{S}_{ti} \tag{7.7} \\
\mathrm{\hat{F}}_{ti}^{(r)} &=\sum_{j}^{t}\hat{f}_{ji}^{(r)} \tag{7.8}
\end{align}




# D Durrant & Müller 2013: Modeling Call Record Data: Examples From Cross-Sectional and Longitudinal Surveys

## 12.3.2.1 Modeling the Outcome at a Particular Call

\begin{align}
\text{log} \left(\frac{\pi_i}{1-\pi_i}\right) &= \beta' x_i \tag{Model 1} \\
\text{log} \left(\frac{\pi_{ij}}{1-\pi_{ij}}\right) &= \beta' x_{ij} +u_j \tag{Model 2}
\end{align}


## 12.3.2.2 Modeling the Outcome Across Calls (Discrete-Time Event History Analysis)

\begin{align}
\text{log} \left(\frac{\pi_{ti}}{1-\pi_{ti}}\right) &= \alpha_t + \beta' x_{ti} \tag{Model 3}
\end{align}


## 12.3.2.3 Modeling the Outcome Across Calls for Repeated Events (Discrete-Time Event History Analysis)

\begin{align}
\text{log}\left(\frac{\pi_{ti}}{1-\pi_{ti}}\right) &= \alpha_t + \beta' x_{ti} + h_i \tag{Model 4}\\
\notag \\
\text{log}\left(\frac{\pi^{(s)}_{tij}}{\pi^{(4)}_{tij}}\right) &= \alpha^{(s)}_{t} + \beta'^{(s)} x^{(s)}_{tij} + h^{(s)}_{ij} + u^{(s)}_j  \tag{Model 5} \\
\notag \\
\text{log}\left(\frac{\pi^{(s)}_{tij}}{\pi^{(4)}_{tij}}\right) &= \alpha^{(s)}_{t} + \beta^{(s)} x^{(s)}_{tij} + \lambda^{(s)} h_{ij} + \gamma^{(s)} u_j  \tag{Model 5$\matr{'}$}
\end{align}







# Literatur



