---
title: "Chapter 8"
output: 
  pdf_document:
    highlight: pygments
    number_sections: true
---



```{r}
library(tidyverse)
library(AzureML)
ws <- workspace()

mlbook_red <- tbl_df(download.datasets(
  dataset = ws, 
  name    = "mlbook2_r.csv"))
# what are the available variables
names(mlbook_red)

# Attach library
library(nlme)
```


Table 8.1 Model 1
```{r}
mod81 <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~IQ_verb|schoolnr, data = mlbook_red, method="ML")
summary(mod81)
```

Model 2
```{r}
mod82 <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~IQ_verb|schoolnr, weights=varIdent(form=~1|sex),
                data = mlbook_red, method="ML")
summary(mod82)
```


nlme offers a number of variance functions that do not include
the linear variance function used in Model 3.
These are treated in Pinheiro & Bates (2004), Section 5.2.
It turns out that the variance functions offered by nlme,
when used in the default way, represent this data set
slightly less well than a linear variance function.

Therefore we work towards the results in Table 8.2
by also using the result obtained by another program (MLwiN),
namely, that the linear variance function for level 1 was estimated as

```{r}
IQ_verb <- mlbook_red$IQ_verb
#36.382 - 2*1.689*IQ_verb
```

We define the transformed variable
```{r}
IQ_trans <- 36.382 - 2*1.689*mlbook_red$IQ_verb
```



We can use this variance function 
(giving it a freely estimated multiplicative parameter, 
which indeed will be estimated very close to 1)
by using varFixed:

```{r}
mod83 <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~ IQ_verb|schoolnr, weights=varFixed(~IQ_trans),
                data = mlbook_red, method="ML")
summary(mod83)
```

Another way is by using varPower
```{r}
mod83a <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~ IQ_verb|schoolnr, weights=varPower(form=~IQ_trans),
                data = mlbook_red, method="ML")
summary(mod83a)
```


and a further additional parameter is allowed by varConstPower
```{r}
mod83b <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~ IQ_verb|schoolnr, 
                weights=varConstPower(form=~IQ_trans),
                data = mlbook_red, method="ML")
summary(mod83b)
```



A more satisfactory approach that does not assume
previous knowledge, is to use a categorized version of IQ_verb:
First, look at the distribution to find reasonable category bounds
```{r}
hist(mlbook_red$IQ_verb)

IQ_cat <- cut(mlbook_red$IQ_verb, c(-10,-2,-1,0,1,2,10))
table(IQ_cat)
```


Now we fit a model where the variance depends on categorized IQ
```{r}
mod83c <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses, 
                random=~ IQ_verb|schoolnr, 
                weights=varIdent(form=~1|IQ_cat),
                data = mlbook_red, method="ML")
summary(mod83c)
```

Parameter const is estimated close to 0 and power close to 0.5,
with the log-likelihood -12215.1; this means 
that effectively the same model results as Model 3 in Table 8.2. 

For Model 4 in Table 8.2, we first compute the transformed variables.
```{r}
IQ_plus2 <- with(mlbook_red, ifelse(IQ_verb > 0, IQ_verb^2, 0))
IQ_min2 <- with(mlbook_red, ifelse(IQ_verb < 0, IQ_verb^2, 0))
```

Have a look at whether this does what we wanted:
```{r}
plot(mlbook_red$IQ_verb, IQ_plus2)
plot(mlbook_red$IQ_verb, IQ_min2)
```


Again working toward the known answer, we transform IQ_verb to
```{r}
IQ_trans2 <- 36.139 - 2*1.769*mlbook_red$IQ_verb 
```


Estimating with varFixed turns out to require some higher
control settings for the algorithm:
```{r}
mod84 <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses + 
                        IQ_plus2 + IQ_min2, 
           random=~ IQ_verb|schoolnr, weights=varFixed(~IQ_trans2),
           control=lmeControl(maxIter=400,msMaxIter=400,msMaxEval=400,
           niterEM=400,nlmStepMax=400,msTol=1E-9,msVerbose=TRUE),
           data = mlbook_red, method="ML")
summary(mod84)
```

Or, again with the categorized version of IQ:
```{r}
mod84c <- lme(langPOST ~ IQ_verb*ses + sex + sch_iqv*sch_ses + 
                        IQ_plus2 + IQ_min2, 
           random=~ IQ_verb|schoolnr, weights=varIdent(form=~1|IQ_cat),
           data = mlbook_red, method="ML")
summary(mod84c)
```



