---
title: "Producing the examples in chapters 4 and 5"
output: 
  pdf_document:
    highlight: pygments
    number_sections: true
---


<!--

###################################################################
###  ch45.r                                                     ###
###                                                             ###
###  This is an R script for producing the examples in          ###
###  chapters 4 and 5 of                                        ###
###  Snijders, Tom A.B., and Bosker, Roel J.                    ###
###  Multilevel Analysis:                                       ###
###  An Introduction to Basic and Advanced Multilevel Modeling, ###
###  second edition.                                            ###
###  London etc.: Sage Publishers, 2012                         ###
###                                                             ###
###  version June 14, 2014                                      ###
###                                                             ###
###################################################################

-->


```{r, include=FALSE}

library(tidyverse)
library(AzureML)
ws <- workspace()

mlbook_red <- tbl_df(download.datasets(
  dataset = ws, 
  name    = "mlbook2_r.csv"))

# what are the names
names(mlbook_red)


# Note that school means can be calculated by functions such as
# schmeans <- aggregate(mlbook_red, by = list(mlbook_red$schoolnr), mean)
# But this is not necessary here because the school means
# are already in the data set.
```



```{r, include=FALSE}
# Attach library
library(lme4)

# How does function lmer work
?lmer
# In the Examples in this help page, data from "sleepstudy" are used.
# This is a longitudinal data set,
# and can serve as an illustration of the type of
# longitudinal data that can be modeled by the Hierarchical Linear Model.
?sleepstudy
```



# Example 4.1
```{r}
mlb0 <- lmer(langPOST ~  (1|schoolnr), data = mlbook_red,
             REML = FALSE)
summary(mlb0)
```


# Example 4.3, Table 4.4

```{r}

mlb44 <- lmer(langPOST ~ IQ_verb + sch_iqv
                        + (1|schoolnr), data = mlbook_red,
                        REML = FALSE)
summary(mlb44)
```



## The parameters of the random part of the model are in
```{r}
VarCorr(mlb44)
```



## the estimated intercept variance is
```{r}
VarCorr(mlb44)$schoolnr[1,1]
```


## For other methods for the objects produced by lmer, see
```{r}
methods(class="merMod")
```


# Section 4.8.
## The posterior means are obtained as follows: the word ranef stands for "random effects"
```{r}
re.mlb44 <- ranef(mlb44, condVar=TRUE, standard=TRUE)
tbl_df(re.mlb44$schoolnr) %>% glimpse
```


## The condVar parameter will also give the posterior variances. What is the structure of this object?
```{r}
str(re.mlb44)
```

##The posterior means are
```{r}
postmean <- re.mlb44$schoolnr[,1]
str(postmean)
```

##  and the posterior variances are
```{r}
postvar <-  attr(re.mlb44$schoolnr,'postVar')[1,1,]
head(postvar)
```



## These are also the comparative variances. The diagnostic variance is calculated using (4.18):
```{r}
diagvar <- VarCorr(mlb44)$schoolnr[1,1] - postvar
diagvar[1:5] 
```


# Comparative standard deviations
```{r}
compsd <- sqrt(postvar)
compsd[1:5]
```


# Bounds of comparative intervals
```{r}
lower <- postmean - 1.39*compsd
upper <- postmean + 1.39*compsd
lower[1:5]
upper[1:5]
```



## Order
```{r}
perm <- order(postmean, lower, upper)
pm_sort <- postmean[perm]
upper_sort <- upper[perm]
lower_sort <- lower[perm]
pm_sort[1:5]
upper_sort[1:5]
pm_sort[1:5]
```



## A caterpillar plot like Fig. 4.4 can be produced as follows.
```{r, eval=FALSE}
library(Hmisc)
errbar(1:211, pm_sort, upper_sort, lower_sort)
```

# Example 5.4

\newpage

```{r}
form <- langPOST ~ 
  1 + 
  IQ_verb + 
  ses + 
  sch_iqv + 
  sch_ses +
  IQ_verb:ses +
  sch_iqv:sch_ses +
  (1 + IQ_verb|schoolnr)

mlb54sh <- lmer(form, data = mlbook_red,
                        REML = FALSE)

summary(mlb54sh)

```

\newpage

```{r}

mlb54 <- lmer(langPOST ~ IQ_verb*ses + sch_iqv*sch_ses
                        + (IQ_verb|schoolnr), data = mlbook_red,
                        REML = FALSE)
summary(mlb54)

```


```{r}
mlb54 <- lmer(langPOST ~ IQ_verb*ses + sch_iqv*sch_ses
                        + (IQ_verb|schoolnr), data = mlbook_red,
                        REML = TRUE)
summary(mlb54)
```





```{r}
library("R2MLwiN")
library("doBy")
options(MLwiN_path = "C:/Program Files (x86)/MLwiN trial/i386/")

form <- langPOST ~ 
  1 + 
  IQ_verb + 
  ses + 
  sch_iqv + 
  sch_ses +
  IQ_verb:ses +
  sch_iqv:sch_ses +
  (1 + IQ_verb|schoolnr)

(mymodel1 <- runMLwiN(form, data = mlbook_red))

```












<!--

########################## nlme ##################################

# The preceding analysis can also be done using nlme:
detach("package:lme4")
library(nlme)
# The main function here is lme
?lme
# and, as you see, the data used in the example of the help page is
?Orthodont
#also a longitudinal data set.

# Example 4.3, Table 4.4
mlb44 <- lme(langPOST ~ IQ_verb + sch_iqv,
                        random = ~ 1|schoolnr, data = mlbook_red, method="ML")
summary(mlb44)

# The estimates of the random part are given by
VarCorr(mlb44)
str(VarCorr(mlb44))
# The estimated intercept variance is
VarCorr(mlb44)$schoolnr[1,1]
# with associated standard deviation
sqrt(as.numeric(VarCorr(mlb44)$schoolnr[1,1]))
# also given by
attr(VarCorr(mlb44)$schoolnr, "stddev")

# To get the standard errors (of which the relevance is minor,
# as they should not be used for testing the hypothesis
# of a zero variance), it may be noted that the
# covariance matrix of the logarithms of the standard deviation parameters
# can be obtained from
(se.log.sd = sqrt(diag(mlb44$apVar)))
# For the component apVar and other components of the object
# returned by lme, look at the help page for
# ?lmeObject

# The standard errors of the intercept variance and corresponding
# standard deviation can be obtained from se.log.var
# by the delta method, see formulae (6.2) in Section 6.3.1.
# Note that se.log.sd = standard error(log(s.d.))
#                     = 0.5*standard error(log(variance))
# From (6.2) we see that the standard error of the variance
# is the variance times the standard error of the log-variance.
# This can be obtained for both the level-2 variance
# and the level-1 variance by
2*se.log.sd * as.numeric(VarCorr(mlb44)[,1])
# The standard errors of the estimated standard deviations are
se.log.sd * as.numeric(VarCorr(mlb44)[,2])

# Example 4.4: The REML estimate
mlb44_reml <- lme(langPOST ~ IQ_verb + sch_iqv,
                        random = ~ 1|schoolnr, data = mlbook_red, method="REML")
summary(mlb44_reml)
# The estimates of the random part using ML are
VarCorr(mlb44)
# and those using REML are
VarCorr(mlb44_reml)

# The posterior means
post.means <- ranef(mlb44_reml)

# Section 4.9
# The data are currently not available.
# A three-level random intercept model can be fitted, e.g., by the command
# model_lev3 <- lme(math ~ IQ, random =list(~1|schoolnr, ~1|classnr))
# if the level-2 identifier is classnr and the level-3 identifier is schoolnr.

# Example 5.4
mlb54 <- lme(langPOST ~ IQ_verb*ses + sch_iqv*sch_ses,
                        random =~ IQ_verb|schoolnr, data = mlbook_red,
                        method="ML")
summary(mlb54)


-->



