---
title: "UserGuide09"
header-includes:
   - \usepackage{bm}
   - \usepackage{amsmath,amssymb,mathrsfs,booktabs}
   - \newcommand{\matr}[1]{\mathbf{#1}} % undergraduate algebra version
author: "Silvan Hüsler"
date: "18 Januar 2017"
output: 
  pdf_document:
    toc: TRUE
    keep_tex: false
    highlight: pygments
    number_sections: false
nocite: | 
  @Hox.2014, @Zhang.2016
# -------------------------------------------------
references:
- id: Hox.2014
  type: chapter
  author:
  - family: Hox
    given: Joop
  - family: Wijngaards-de Meij
    given: Leoniek
  editor:
  - family: Best
    given: Henning
  - family: Wolf
    given: Christof
  issued:
  - year: '2014'
  title: The multilevel regression model
  container-title: The sage handbook of regression analysis and causal inference
  publisher: SAGE Publications Ltd
  page: '133-152'
  DOI: 10.4135/9781446288146.n7
  ISBN: '9781446252444'
# -------------------------------------------------
- id: Zhang.2016
  type: article-journal
  author:
  - family: Zhang
    given: Zhengzheng
  - family: Parker
    given: Richard M. A.
  - family: Charlton
    given: Christopher M. J.
  - family: Leckie
    given: George
  - family: Browne
    given: William J.
  issued:
  - year: '2016'
  title: 'R2MLwiN: A package to run mlwin from within r'
  title-short: R2MLwiN
  container-title: Journal of Statistical Software
  volume: '72'
  issue: '10'
  DOI: 10.18637/jss.v072.i10
  ISSN: '1548-7660'
# *+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```


```{r pakete, include=FALSE}
library("tidyverse")
library("purrr")
library("lazyeval")
library("readr")
library("haven")
library("lubridate")
library("magrittr")
library("stringr")
library("tibble")
library("zoo")
library("dygraphs")
library("plotly")
library("glmmML")
library("survival")
library("forcats")
library("wesanderson")
library("car")
options(tibble.print_min = 100)
options(scipens = 9999999999999999999)
library("R2MLwiN")
library("doBy")
options(MLwiN_path = "C:/Program Files (x86)/MLwiN trial/i386/")
```


## 9.1 Introduction and description of the example data 

```{r}
data(bang, package = "R2MLwiN")
summary(bang)
```



## 9.2 Single-level logistic regression

### Link functions 

### Interpretation of coeficients

### Fitting a single-level logit model in MLwiN

```{r}
addmargins(with(bang, table(lc, use)))
```


```{r}
(mymodel1 <- runMLwiN(logit(use) ~ 1 + lc, D = "Binomial", data = bang))
```


```{r}
linearHypothesis(mymodel1, "FP_lcOne_child = FP_lcTwo_children")
```


### A probit model

```{r}
(mymodel2 <- runMLwiN(probit(use) ~ 1 + lc, D = "Binomial", data = bang))
```


```{r}
(mymodel3 <- runMLwiN(logit(use) ~ 1 + lc + age, D = "Binomial", data = bang))
```

## 9.3 A two-level random intercept model 

### Model specification

### Estimation procedures

### Fitting a two-level random intercept model in MLwiN

```{r}
(mymodel4 <- runMLwiN(
  logit(use) ~ 1 + lc + age + (1 | district), 
  D = "Binomial", 
  data = bang))
```


```{r}
(mymodel5 <- runMLwiN(
  logit(use) ~ 1 + lc + age + (1 | district), 
  D = "Binomial", 
  estoptions = list(
    nonlinear = c(N = 1,M = 2), 
    startval = list(
      FP.b = mymodel4@FP, 
      FP.v = mymodel4@FP.cov, 
      RP.b = mymodel4@RP, 
      RP.v = mymodel4@RP.cov)),
  data = bang))
```


```{r}
linearHypothesis(mymodel5, "RP2_var_Intercept = 0")
```


### Variance partition coeficient

```{r}
set.seed(1)

invlogit <- function(x) exp(x)/(1 + exp(x))

u <- sqrt(coef(mymodel5)["RP2_var_Intercept"]) * qnorm(runif(5000))

p1 <- invlogit(coef(mymodel5)["FP_Intercept"] + u)

p2 <- invlogit(coef(mymodel5)["FP_Intercept"] + 
                 coef(mymodel5)["FP_lc3plus"] + 
                 coef(mymodel5)["FP_age"] * -9.7 + u)

p3 <- invlogit(coef(mymodel5)["FP_Intercept"] + 
                 coef(mymodel5)["FP_age"] * 15.3 + u)

v1 <- p1 * (1 - p1)
lev2var1 <- sd(p1)^2
lev1var1 <- mean(v1)

v2 <- p2 * (1 - p2)
lev2var2 <- sd(p2)^2
lev1var2 <- mean(v2)

v3 <- p3 * (1 - p3)
lev2var3 <- sd(p3)^2
lev1var3 <- mean(v3)

cat(
  paste0("VPC = ", lev2var1/(lev2var1 + lev1var1)))

cat(
  paste0(
    "VPC for a young women with 3+ children (low probability use) = ", 
    lev2var2/(lev2var2 + lev1var2)))

cat(
  paste0(
    "VPC for an old woman with no children (high probability use) = ", 
    lev2var3/(lev2var3 + lev1var3)))
```


### Adding further explanatory variables

```{r}
table(bang$educ)

(mymodel6 <- 
    runMLwiN(
      logit(use) ~ 1 + lc + age + urban + educ + hindu + (1 | district), 
      D = "Binomial", 
      estoptions = list(
        nonlinear = c(N = 1, M = 2), 
        startval = list(
          FP.b = mymodel5@FP, 
          FP.v = mymodel5@FP.cov, 
          RP.b = mymodel5@RP, 
          RP.v = mymodel5@RP.cov)),
  data = bang))
```


## 9.4 A two-level random coeficient model

```{r}
(mymodel7 <- 
   runMLwiN(
     logit(use) ~ 1 + lc + age + urban + educ + hindu + 
       (1 + urban | district), 
     D = "Binomial", 
     estoptions = list(
       nonlinear = c(N = 1, M = 2), 
       startval = list(FP.b = mymodel6@FP, 
                       FP.v = mymodel6@FP.cov, 
                       RP.b = mymodel6@RP, 
                       RP.v = mymodel6@RP.cov)), 
     data = bang))
```

```{r}
linearHypothesis(
  mymodel7, 
  "RP2_cov_Intercept_urbanUrban = 0")

linearHypothesis(
  mymodel7, 
  "RP2_var_urbanUrban = 0")

linearHypothesis(
  mymodel7, 
  c("RP2_cov_Intercept_urbanUrban = 0", "RP2_var_urbanUrban = 0"))
```


```{r}
(mymodel8 <- 
   runMLwiN(
     logit(use) ~ 1 + lc + age + urban + educ + hindu + d_lit + d_pray + 
       (1 + urban | district),
  D = "Binomial", 
  estoptions = list(
    nonlinear = c(N = 1, M = 2), 
    startval = list(
      FP.b = mymodel7@FP, 
      FP.v = mymodel7@FP.cov,
      RP.b = mymodel7@RP, 
      RP.v = mymodel7@RP.cov)), 
  data = bang))
```


## 9.5 Modelling binomial data

### Modelling district-level variation with district-level proportions

### Creating a district-level data set

```{r}
bangshort <- summaryBy(
  use + cons ~ district + d_lit + d_pray, 
  FUN = c(mean, sum), data = bang)

bangshort$use.sum <- NULL
colnames(bangshort) <- c("district", "d_lit", "d_pray", "use", "cons", "denom")
bangshort$use <- bangshort$use - 1
```


### Fitting the model

```{r}
(mymodel9 <- 
   runMLwiN(
     logit(use, denom) ~ 1 + d_lit + d_pray + (1 | district), 
     D = "Binomial", data = bangshort))
```


```{r}
(mymodel10 <-
   runMLwiN(
     logit(use, denom) ~ 1 + d_lit + d_pray + (1 | district), 
     D = "Binomial", 
     estoptions = list(
       nonlinear = c(N = 1, M = 2), 
       startval = list(
         FP.b = mymodel9@FP, 
         FP.v = mymodel9@FP.cov, 
         RP.b = mymodel9@RP, 
         RP.v = mymodel9@RP.cov)),
     data = bangshort))
```




## Literatur



